# OMPutils

The concept behind this package is to implement a variety of base R functions in C++ using Rcpp. It is further intended that these functions will be implemented with implicint parallelism, using either OpenMP or RcppParallel. This package is inspired on two fronts, 

Firstly there are base R functions that can be improved to reduce computation times. For instance, the current implementation of the `rank()` funciton is not very efficient on numeric vectors. If we implement this funciton in Rcpp, we can achieve the following:

```{r}
library(OMPutils)
library(microbenchmark)
x <- rnorm(1E4)

microbenchmark(rank(x), OMPutils:::rcpp_rank(x))
```

Further, most R functions are not parallel, i.e. they are only able to run on a single CPU. Modern computers are have multiple CPUs and most of these cores are often not exploited by R functions. While it is well known that performance gains are not linear with respect to the number of cores used, processing times can be greatly decresed. For instance:

```{r}
library(OMPutils)
library(microbenchmark)
x <- rnorm(1E8)

microbenchmark(sum(x), OMPutils:::rcpp_sum(x),OMPutils:::rcpp_sum(x,threads = 4))
```

